---
title: "Creating graphs with ggplot2"
author: "Eileen"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot 2

We are going to make plots in R using ggplot2 package from the tidyverse


We will install these packages: 

- tidyverse
- readxl
- here

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")
```

# California National Parks Data

```{r read-data}
ca_np <- read_csv(here("data/ca_np.csv"))
# also be written
ca_np <- read_csv("data/ca_np.csv")
ci_np <- read_excel(here("data","ci_np.xlsx"))

```

### Make my own first graph! 

We're going to make a line plot of year (x-variable) versus visitor (y-variable) in ggplot: 

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point()


# Update the geom_ to geom_point(), geom_col(), geom_area()

```

Make a base layer for my CI plots: 
```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))
gg_base + geom_point()
```

### Now let's start customizing

- Line / point colors:  color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed",
  )
```

Create a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch) of a non-default size (size = #) and that are an updated color.  

```{r}
gg_base + 
  geom_point(
    pch = 17,
    size = 4,
    color = "dodgerblue"
  )
```

### Update graph aesthetics based on variables
# Add an arguement to update the color based on the year value
```{r}
gg_base + 
  geom_point(
    aes(size = visitors,
        color = year),
    pch = 10
  )
```




